### 2007-03-15 wikiplia/paper/wikiplia.tex
start paper

\documentclass{article}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}

\begin{document}

\title{{\bf \huge Wikiplia}:\\
       The Free Programming Language\\
       That Anyone Can Edit}

\author{Tom Murphy VII}

\maketitle

\begin{abstract}
We present a new programming language called Wikiplia. We argue that
this language is Free to a greater extent and in more dimensions than
any existing language.

\end{abstract}

\nocite{alice}

{\small
\bibliographystyle{plain}
\bibliography{wikiplia}
}

\end{document}