### 2007-03-16 trunk/wikiplia/paper/wikiplia.tex
syntax

\documentclass[twocolumn]{article}
\usepackage{fullpage}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{proof}

\begin{document}

\newcommand\wcite[1]{\footnote{\tiny Wikipedia, the free encyclopedia: {\it #1}; 2007}}
\newcommand\comment[1]{}
\newcommand\z{\ensuremath{\!}}

\newcommand\D\Delta
\newcommand\G\Gamma
\newcommand\m\mapsto
\newcommand\eval{{\sf eval\,\,}}
\newcommand\tag[2]{{\tt{\mbox{\tt <}}{#1}{\mbox{\tt>}}}{#2}{\tt{\mbox{\tt</}}{#1}{\mbox{\tt>}}}}
\newcommand\prim{{\,\,\sf prim}}
\newcommand\rate{{\sf rate\,\,}}

\newcommand\lb{\ensuremath{[\![}}
\newcommand\rb{\ensuremath{]\!]}}

\title{{\bf \huge Wikiplia}:\\
       The Free Programming Language\\
       That Anyone Can Edit}

\author{Tom Murphy VII}
\date{1 April 2007}

\maketitle

\begin{abstract}
We present a new programming language called Wikiplia. The language
has an unprecedented level of integration: The system is its own
compiler, language definition, documentation, development environment,
distributed filesystem, database, revision control system, software
license, community message board, and World Wide Web home site.
Wikiplia is designed to be Free to a greater extent and in more
dimensions than existing languages.
\end{abstract}

\section{Introduction}

One of the most cherished social principles of mankind is
freedom,\z\wcite{Freedom} in its many incarnations. More recently,
freedom has become an important principle in computer science as well,
with the introduction of Free Software licenses such as the GNU
GPL,\z\wcite{GNU General Public License} extensible markup languages
such as XML,\z\wcite{XML} the ability expicitly deallocate memory
with the {\tt free(3)} library call, and the widespread availibility of
Free Herbal V1agara on the World Wide Web.\z\wcite{Sildenafil}

The aim of this project is to develop a programming language that is
as free as possible. We begin by enumerating freedoms that we desire
to support. Because freedom is a possession of inestimable
value\wcite{Cicero} we do not attempt to rank these freedoms; instead,
each freedom is ``numbered'' using a symbol drawn from incomparable
sets of glyphs.

\paragraph{Freedom $\alpha$: The freedom to tinker.}
Users should be able to study a program to see how it works, and to
make modifications to suit his or her needs. For most software, this
means that the programmer needs access to the software's
documentation, source code, and UML\wcite{Unified Modeling Language}
use case diagrams. This is traditionally achieved through licenses
such as the GPL; however, as we will discuss in
Section~\ref{sec:bootstrap} there are special considerations for
bootstrapping compilers that render the GPL inadequate for this
purpose.

% freedom of speech?

\paragraph{Freedom $\Game$: Freedom of beer.}
Users should be able to write software without paying money to a
licensing authority or certification program.

\paragraph{Freedom $\surd$: Freedom of USA \#1.}
Wikiplia is 100\% made in the USA and only available in
English.\z\wcite{Freedom fries}

% XXX more...

\section{Reflections on booting boots} \label{sec:bootstrap}

\section{Core calculus} \label{sec:calculus}

\begin{figure}[ht]

\comment{
    List of exp list       <list> ... </list>
  | String of string       <string>s</string>
  | Int of IntInf.int      <int>i</int>
  | Symbol of string       <symbol>s</symbol>
  (* the only lazy expression *)
  | Quote of exp           <quote>exp</quote>
  (* user can't write these *)
  | Prim of prim           <prim>p</prim>
  | Closure of (string * exp) list * string * exp
}
\begin{center}
\begin{tabular}{rcl}
 {\em X} & ::= & {\tt<}{\em tag}{\tt >} $X_1\ X_2\ \ldots\ X_n$ {\tt</}{\em tag}{\tt >} \\
         & $|$ & string
\end{tabular}
\end{center}
\caption{Syntax of XML} \label{fig:xml}
\end{figure}

\begin{figure*}[ht]
\[\begin{array}{cc}
% basic, easy
\infer{\G \vdash \eval \tag{string}{s} \m \tag{string}{s}}{} &
\infer{\G \vdash \eval \tag{quote}{X} \m X}{} \\[1em]
\infer{\G \vdash \eval \tag{int}{s} \m \tag{int}{s}}{} &
\infer{\G \vdash \eval \tag{prim}{s} \m \tag{prim}{s}}{} \\[1em]
\infer{\G \vdash \eval \tag{symbol}{s} \m \tag{prim}{s}}{s \prim} &
\infer{\G \vdash \eval \tag{symbol}{s} \m X}{\G(s) = X} \\[1em]
\multicolumn{2}{c}{\infer{\G \vdash \eval \tag{closure}{\G\ s\ X} \m \tag{closure}{\G\ s\ X}}{}} \\[1em]
\multicolumn{2}{c}{\infer{\G \vdash \eval \tag{list}{X_1\ \ldots\ X_n} \m X'}%
{\G\vdash \eval X_1 \m X_1' & \cdots & \G\vdash \eval X_n \m X_n' &
 \G\vdash \rate X_1'\ \ldots\ X_n' \m X'
}} \\[1em]
% todo:
% skip? abort, handle

%
\end{array}\]

\caption{Evaluation of XML, part 1. The judgment $\G \vdash \eval X \m X'$
indicates an assessment of the document $X$ with value $X'$. The
judgment $\rate$ is an auxiliary assessment of a list of documents. It is defined in Figure~\ref{fig:xmlrate}.
$\vec{X}$ is shorthand for a possibly empty sequence of XML documents.
$\G$ is itself an XML document of the form
% XXX is this right wrt <symbol>?
$\tag{list}{\tag{symbol}{s_1}\ X_1\ \ldots\ \tag{symbol}{s_n}\ X_n}$.
We take the judgment $\G(s) = X$ to produce the leftmost $X_i$ in $\G$
such that $s_i$ is $s$. $\G, s = X$ is $\tag{list}{\tag{symbol}{s}\ X
\vec{X}}$ if $\G$ is $\tag{list}{\vec{X}}$. The judgment $s\prim$ holds when
$s$ is one of 
{\sf insert},
{\sf head},
{\sf read},
{\sf abort},
{\sf lambda},
{\sf list},
{\sf cons},
{\sf quote},
{\sf string},
{\sf xcase},
{\sf size},
{\sf sub},
{\sf substr},
{\sf handle},
{\sf parse},
{\sf eval},
{\sf eq},
{\sf +},
{\sf -},
{\sf int},
{\sf history},
{\sf let}, or
{\sf if}.
%
} \label{fig:xmleval}
\end{figure*}

\begin{figure*}[htp]
\[\begin{array}{cc}
% evaluation of lists
\multicolumn{2}{c}{\infer{\G \vdash \rate \tag{prim}{\sf list}\ \vec{X} \m \tag{list}{\vec{X}}}{}} \\[1em]
\multicolumn{2}{c}{\infer{\G \vdash \rate \tag{prim}{\sf cons}\ X\ \tag{list}{\vec{X}} \m \tag{list}{X\ \vec{X}}}{}} \\[1em]
\multicolumn{2}{c}{\infer{\G \vdash \rate \tag{prim}{\sf lambda}\ \tag{symbol}{s}\ X \m \tag{closure}{\G\ s\ X}}{}} \\[1em]
\infer{\G \vdash \rate \tag{closure}{\G'\ s\ X}\ \vec{X} \m X'}{\G',s = \tag{list}{\vec{X}} \vdash \eval X \m X'} &
\infer{\G \vdash \rate \tag{prim}{\sf eval}\ X \m X'}{\G \vdash \eval X \m X'} \\[1em]
\multicolumn{2}{c}{\infer{\G \vdash \rate \tag{prim}{\sf xcase}\ \tag{list}{}\ X\ \vec{X} \m X'}{\G \vdash \eval X \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{list}{X_h \vec{X_t}}\  X_0\ \\
                  & \tag{list}{\tag{symbol}{s_h}\ \tag{symbol}{s_t}\ X_b}\ \vec{X} \m X' \\
\end{array}
}{\G, s_h = X_h, s_t = X_t \vdash \eval X_b \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{quote}{X}\  X_0\ X_1\ \\
                  & \tag{list}{\tag{symbol}{s_q}\ X_b}\ \vec{X} \m X' \\
\end{array}
}{\G, s_q = X \vdash \eval X_b \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{string}{s}\  X_0\ X_1\ X_2\ 
                    X\ \vec{X} \m X' \\
\end{array}
}{\G \vdash \eval X \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{int}{s}\  X_0\ X_1\ X_2\ X_3\ 
                    X\ \vec{X} \m X' \\
\end{array}
}{\G \vdash \eval X \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{symbol}{s}\  X_0\ X_1\ X_2\ X_3\ X_4\ \\
                  & \tag{list}{\tag{symbol}{s_s}\ X_b}\ \vec{X} \m X' \\
\end{array}
}{\G, s_s = \tag{string}{s} \vdash \eval X_b \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{t}{\vec{X_t}}\  X_0\ X_1\ X_2\ X_3\ X_4\ X_5\ X \vec{X} \m X' \\
\end{array}
}{t = {\tt prim} \,\,or\,\, {\tt closure} & \G \vdash \eval X \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{\G \vdash \rate \tag{prim}{\sf quote}\ {X} \m \tag{list}{X}}{}} \\[1em]
%
\multicolumn{2}{c}{\infer{
\G \vdash \rate \tag{prim}{\sf head}\ \tag{string}{s} \m X
}{{\tt cvs\ checkout}\ s = X}} \\[1em]
\multicolumn{2}{c}{\infer{
\G \vdash \rate \tag{prim}{\sf read}\ \tag{string}{s} \tag{int}{i} \m X
}{{\tt cvs\ checkout\ -\!r}\ i\ s = X}} \\[1em]
\multicolumn{2}{c}{\infer{
\G \vdash \rate \tag{prim}{\sf insert}\ \tag{string}{s} X \m \tag{int}{i}
}{{\tt cvs\ commit}\ s\ X = i}} \\[1em]
\multicolumn{2}{c}{\infer{
\G \vdash \rate \tag{prim}{\sf history}\ \tag{string}{s} \m \tag{list}{\vec{X}}
}{{\tt cvs\ log}\ s = \vec{X}}} \\[1em]

\end{array}\]

\caption{Evaluation of XML, part 2. The {\sf rate} judgment assesses a
sequence of XML expressions. The rules for rating the primitives {\sf
string}, {\sf sub}, {\sf substr}, {\sf +}, {\sf -}, {\sf int}, {\sf
eq} and {\sf if} are omitted for space.
%
} \label{fig:xmlrate}
\end{figure*}

Wikiplia is built upon a core calculus of structured data with
primitive revision control. Because we wish to support the freedom to
tinker, the structured data take the form of XML (the {\em extensible}
markup language; Figure~\ref{fig:xml}). Similar to the W3C's XML
Validation,\z\wcite{XML schema} we allow the quality of an XML
document to be assessed via a process called {\em evaluation}, whose
output (if any) is itself an XML document. Selected rules for XML
evaluation (the dynamic semantic markup) are given in
Figures~\ref{fig:xmleval}~and~\ref{fig:xmlrate}.\z\footnote{XML
documents can be self-correcting through the use of the {\sf handle}
primitive, which detects an invalid document and proceeds along an
alternative path. We omit the rules for this feature, which requires
propagating invalid document status throughout evaluation and thus
complicates the rules substantially.}

Revision control is accessed through the class of imperative {\tt cvs}
judgments. We assume a single global repository, which maps {\em keys}
(strings) to lists of {\em revisions}. A revision is a monotonically
increasing and unique {\em revision number} (integer) paired with an
XML document. The judgment ${\tt cvs commit}\ s\ X = i$ creates a new
revision with revision number $i$ and data $X$ and inserts it under
the key $s$. The judgment ${\tt cvs checkout}\ s = X$ fetches the most
recent revision for the key $s$, the document $X$ (if no such key
exists, then the document is invalid). The judgment ${\tt cvs\
checkout\ -\!r}\ i\ s = X$ does the same, but for the specific
revision number $i$ (if no such revision exists, the document is
invalid). Finally, ${\tt cvs\ log}\ s = \vec{X}$ fetches all of the
revision numbers for the key $s$, as a series of integers $\vec{X}$.

\subsection{Syntax}

\begin{figure}[b]
\begin{center}
\begin{tabular}{c}
  \begin{tabular}{rcl}
   {\em E} & ::= & {\tt(} $E_1\ E_2\ \ldots\ E_n$ {\tt)} \\
           & $|$ & {\tt"}string{\tt"} \\
           & $|$ & $n$ \\
           & $|$ & symbol \\
           & $|$ & {\tt'}E \\
  \end{tabular} \\[3em]

  \begin{tabular}{r@{\,\,=\,\,}l}
   \lb {\tt(} $E_1\ \ldots\ E_n$ {\tt)}\rb & $\tag{list}{ \lb E_1 \rb\ \ldots\ \lb E_n \rb}$ \\
   \lb {\tt"}$s${\tt"}\rb & $\tag{string}{s}$ \\
   \lb $n$\rb & $\tag{int}{n}$ \\
   \lb $sym$\rb & $\tag{symbol}{sym}$ \\
   \lb {\tt'}$E$\rb & $\tag{quote}{E}$ \\
  \end{tabular}

\end{tabular}
\end{center}
\caption{The syntax for the XESP syntax. The recursively defined $\lb
\cdot \rb$ operation converts an XESP expression into an XML
document.} \label{fig:xesp}
\end{figure}

XML documents are universally parseable.\z\wcite{Parsing} However,
they are difficult to write and read. Therefore, as usual\wcite{RELAX
NG} we create a new syntax by which humans can write and read programs
and which the computer automatically parses and converts to the easily
parsable XML syntax and back to the new syntax, reducing
complexity.\z\footnote{For efficiency, the Wikiplia implementation is
optimized to lazily perform these translations, so that the document
is never represented in XML form.} This compact syntax is based upon
parentheses rather than tags: The XML document $\tag{list}{X_1\ X_2}$
is instead written ${\tt(}X_1\ X_2{\tt)}$. Note that the closing
parenthesis is not {\em named} as in XML, which makes parsing
difficult because the computer must {\em guess} which parenthesis
belongs to which other parenthesis. We therefore call this compact
syntax XESP because it basically {\em reads the programmer's mind} to
guess what the name of the closing parenthesis should be. The grammar
for XESP is given in Figure~\ref{fig:xesp} along with the translation
to XML documents.

\end{document}