### 2011-03-05 wishlist/paper.tex
commit so I can profile on linux

facebook wall posts, 4-grams
100 top paths:
0.002525504 pittsburgh
0.002093510 steelers
0.002093510 sfo
0.001953943 it's
0.001256106 bdl
0.001116539 sigbovik
0.001099196 facebook
0.000976971 mic
0.000976971 s
0.000976971 can't
0.000837404 i'm
0.000837404 icfp
0.000837404 app
0.000697837 x
0.000697837 drunj
0.000697837 g
0.000616021 ther
0.000558269 doesn't
0.000558269 's
0.000558269 sr
0.000558269 pm
0.000558269 fyi
0.000558269 wtf
0.000558269 k
0.000558269 don't
0.000558269 today's
0.000558269 m
0.000558269 what's
0.000456766 iphone
0.000429438 inted
0.000418702 oh
0.000418702 phl
0.000418702 gonna
0.000418702 pre
0.000418702 zrh
0.000418702 th
0.000418702 h
0.000418702 b
0.000418702 n
0.000418702 ok
0.000418702 d
0.000418702 that's
0.000418702 rex
0.000418702 fanzibar
0.000418702 st
0.000418702 kinda
0.000418702 didn't
0.000355262 cally
0.000334962 brillobox
0.000314027 superburrito
0.000305304 weath
0.000279135 techno
0.000279135 vii

trigrams, including comments:
100 top paths:
0.003013897 it's
0.001856422 ther
0.001704432 i'm
0.001503910 sfo
0.001102867 -
0.001102867 oh
0.001050350 don't
0.001002607 haha
0.001002607 s
0.000902346 bdl
0.000808769 mic
0.000808769 that's
0.000776981 pittsburgh
0.000705812 thes
0.000701825 ok
0.000645063 can't
0.000601564 icfp
0.000601564 k
0.000583335 didn't
0.000507201 runj
0.000501303 g
0.000501303 omg
0.000501303 x
0.000485262 ins
0.000453126 sout
0.000450580 ands
0.000449950 coff
0.000425932 reat
0.000425133 ning
0.000416189 pittle
0.000412468 getty
0.000407151 ster
0.000401043 sr
0.000401043 pm
0.000401043 m
0.000401043 wtf
0.000401043 aw
0.000401043 d
0.000401043 fyi
0.000401043 ple
0.000401043 's
0.000400791 phon
0.000386315 weat
0.000384116 befor
0.000375978 els
0.000374307 als
0.000357782 nothe
0.000354769 app
0.000354769 appy
0.000354422 they're
0.000353750 wate
0.000345727 there's
0.000340508 res
0.000339344 mes
0.000334202 ove
0.000326473 wher
0.000324373 lity
0.000319011 hotos
0.000316613 dea
0.000315777 sigbovik
0.000315105 ver
0.000314965 thand
0.000310607 outh
0.000300782 n

facebook, n = 2
100 top paths:
0.010628671 ing
0.007144197 th
0.005025442 st
0.004291375 re
0.002932937 ne
0.002662094 se
0.002546194 whe
0.002108786 ther
0.001923496 al
0.001669458 con
0.001652820 wor
0.001498222 ch
0.001495485 wis
0.001379778 i'm
0.001336166 jus
0.001302469 sh
0.001280888 yout
0.001244352 fo
0.001225250 le
0.001224422 pre
0.001221357 ar
0.001201881 com
0.001200574 ge
0.001165417 thes
0.001116295 som
0.001108305 fing
0.001104851 ang
0.001102867 -
0.001102706 it's
0.001039704 ist
0.001036886 hing
0.001007930 ou
0.001002607 s
0.000967282 de
0.000932965 en
0.000926568 tor
0.000827260 wat
0.000827151 oh
0.000814235 cor
0.000812961 ned
0.000783696 lis
0.000762020 ple
0.000760162 whis
0.000759898 sed
0.000755396 res
0.000728412 mor
0.000728049 ong
0.000727818 don
0.000720098 cand
0.000710277 ning
0.000708380 ithe
0.000693547 ame
0.000659248 ty
0.000656211 dow
0.000635771 ha
0.000633225 ext
0.000601564 k