### 2012-03-10 pants/pants.tex
some progress on this fine paper

\documentclass[twocolumn]{article}
\usepackage{graphicx}
\usepackage{amssymb}

\begin{document}

\newcommand\comment[1]{}

\title{{\bf \huge i-PHONE app stor }:\\
       where is my pants
       }

\author{Dr. Tom Murphy VII, Ph.D. $\qquad$ Dr. Abigale G. Lade, Ph.D.}
\date{1 April 2012}

\maketitle

\begin{abstract}
Dear I--PHONE app stor\,\,\,\, please accept my applicatian for publishment of my game
\end{abstract}

\footnotetext[-1]{Copyright \copyright\ 2012 the Regents of the Wikiplia
Foundation. Appears in SIGBOVIK 2012 with the permission of the
Association for Computational Heresy; {\em IEEEEEE!} press,
Verlag-Verlag volume no.~0x40-2A. Distribute free-ily. $\stackrel\Game\heartsuit$ 0.00}

\vspace{1em}
{\noindent \small {\bf Keywords}:
  pants, where, is my; juvenilia
}

\section{Introduction}

Dear {\sf app stor}$\quad$ I want you to have my game . My game is
{\it where is my pants}. Where is my Pants is a newwest game for the i-touch (and I-PAD, ect.). I believe it should be publiced it the {\sf app stor}. For every1 to play the game and find my pants. Because

\subsection{Reasons because it should be publiced}
\begin{enumerate}
\item This game it SO FUN!!
\item $\cup$ can play it any time or any place
\item I believe that if you put your mind to it, $\cup$ can accomplish anything
\item There are one thousand people with i-touch (and !`-PAD ect.) . If evry touch and/or PAD bought this game. Then I would have at least RICH!!!!!!!
\item I don't know where did my pants went
\end{enumerate}

\subsection{Price of game}

\begin{enumerate}
\item The price of the {\it where is my pants} will be \$9.99 USA.
\end{enumerate}

\subsection{How to play}

\begin{enumerate}
\item get down load game in i- touch
\item lunch game
\item look in the screen to see if you find pants
\item touch pants to see where is my pants
\item lotta of people have pants! so keep touching
\item if the pants 
\end{enumerate}

\end{document}