### 2014-02-22 trunk/epsilon/paper.tex
cxpoint to hacksession

\documentclass[twocolumn]{article}
\usepackage[utf8]{inputenc}
\usepackage[top=1.1in, left=0.85in, right=0.85in]{geometry}

% \usepackage{eclbkbox}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{code}
% \usepackage{amscd}
% \usepackage{xy}
\usepackage{graphicx}

% \pagestyle{empty}

% \usepackage{ulem}
% go back to italics for emphasis, though
% \normalem

\usepackage{natbib}

\begin{document}

\title{What, if anything, is epsilon?}
\author{Dr.~Tom~Murphy~VII~Ph.D.\thanks{
Copyright \copyright\ 2013 the Regents of the Wikiplia
Foundation. Appears in SIGBOVIK 2013 with the careless accounting of the
Association for Computational Heresy; {\em IEEEEEE!} press,
Verlag-Verlag volume no.~0x40-2A.
\yen 0.00}
}

\renewcommand\>{$>$}
\newcommand\<{$<$}

\date{1 April 2014}

\maketitle

\begin{abstract}
what do I put here
\end{abstract}

\vspace{1em}
{\noindent \small {\bf Keywords}:
 computational archaeology, epsilon, very-small and medium-small numbers
}

\section*{Introduction}

Epsilon, the all-spelled-out version of $\epsilon$, although properly
``epsilon'' because it's the {\em lowercase} Greek letter, though it's
not like I'm going to start my paper with a lowercase letter even if
it's technically correct, since I like to wait at least until the
second or third letter of the paper before the reader starts doubting
that I can write or spell or have shift-keys on my keyboard, anyway
epsilon is a mathematical symbol denoting a {\em very small number}.

% (intro about math definition in \epsilon--\delta formulation of limits.
% it's not that complicated)

In mathematics, $\epsilon$ usually refers to the
$\epsilon--\delta$ formulation of limits. This is pretty simple and a
reminder appears in Figure~\ref{fig:epsilondelta}. This paper is not
about that kind of math.

In computing, $\epsilon$ is used in a much more general sense to just
mean some small number or error bound. For example, two numbers are
often considered equal if their absolute difference is less than
$\epsilon$.

In IEEE-754 floating-point~\cite{ieee754}, the standard way that
computers represent ``real numbers,'' there is a specific formal value
called ``machine epsilon'', or ``unit roundoff''. It is the maximum
(relative) amount of error from a single rounding operation. This
number is useful if you want to do numerical programming and be
careful about what you're doing.

Most programmers find this subject too tedious and simply pick a
number that seems pretty small. Thus in practice, $\epsilon$ is an
application-specific choice. Of well-known ``constants,''
$\epsilon$ may be the least agreed-upon, with values seen in the wild
spanning {\em 50 orders of magnitude}. This paper explores the practical
values of $\epsilon$ in real software.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.90 \linewidth]{epsilondelta.pdf}
\end{center}\vspace{-0.1in}
\caption{explain the graph}
\label{fig:epsilondelta}
\end{figure}

\section{Methodology}

Did you know that the SPEC benchmarks~\cite{spec2006} cost \$800? Like
they literally expect me to pay them money to download the source code
so that I could grep for {\tt const double epsilon} or test my
compiler out on that. Mmany are even based on open-source software like
Sphinx and POV-Ray. Ridiculous. I refuse. Values of epsilon for the
SPEC benchmarks do not appear in Figure~\ref{fig:spec}.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.99 \linewidth]{spec.pdf}
\end{center}\vspace{-0.1in}
\caption{\$800? Fuck that!}
\label{fig:spec}
\end{figure}

\section{Results}

\bibliography{paper}{}
\bibliographystyle{plain}
\end{document}