### 2014-02-22 chesstego/paper.tex
Create Chesstego with search

coauthors:

Jim McCann, in italics ...... oh! .... nervous laughter .... no, that's not part of the name still. I stopped saying the name. Why are you still typing? This isn't the name any more. End quote.