### 2014-03-01 chesstego/paper.tex
Add COPYING to epsilon

\documentclass[twocolumn]{article}
\usepackage[top=1.1in, left=0.85in, right=0.85in]{geometry}

\usepackage{url}
% \usepackage{code}
% \usepackage{cite}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{chessboard}

\pagestyle{empty}

\usepackage{ulem}
% go back to italics for emphasis, though
\normalem

\usepackage{natbib}

% \newcommand\comment[1]{}
\newcommand\sfrac[2]{\!{}\,^{#1}\!/{}\!_{#2}}

\begin{document}

\title{It still seems that black has hope in these extremely unfair variants of chess}
\author{Dr.~Tom~Murphy~VII~Ph.D. \and Ben Blum \and
Jim McCann, in italics ...... oh! .... nervous laughter .... no, that's not part of the name still. I stopped saying the name. Why are you still typing? This isn't the name any more. End quote.\thanks{
Copyright \copyright\ 2014 the Regents of the Wikiplia
Foundation. Appears in SIGBOVIK 2014 with the undivided
attention of the Association for Computational Heresy; 
{\em IEEEEEE!} press, Verlag-Verlag volume no.~0x40-2A.
\$0.00}}

\renewcommand\th{\ensuremath{{}^{\textrm{th}}}}
\newcommand\st{\ensuremath{{}^{\textrm{st}}}}
\newcommand\rd{\ensuremath{{}^{\textrm{rd}}}}
\newcommand\nd{\ensuremath{{}^{\textrm{nd}}}}
\newcommand\at{\ensuremath{\scriptstyle @}}

\date{1 April 2014}

\maketitle \thispagestyle{empty}

\begin{abstract}
what do I put here
\end{abstract}

\setchessboard{tinyboard,showmover=false}
% \chessboard[
%   setfen=rnbqkbnr/pppppppp/8/8/8/8/NNNBBNNN/QQQQKQQQ]

\input{king2}

% results:

% No mate in 12 for 0-7. 6 ran for like 2 full days!
% regular board.

% No mate in 14 for 1,2,3,4,5,7. didn't run 7 because 6 is
% still running overnight! 13h51m! 0 also running for 42h00m.

% No mate in 10 for king at pawn index 0-7, even when
% white's pieces are all queens (but the king)
% (6 also finished last. Is this the weakest spot?)

% Now mate in 10 for white having ``mega'' pieces

% Mate in 10 for pawn idx 0! 441450 tree 
%                 and idx 1! 210316 tree

% TODO: Generate ``book'' where white is making strong moves
% (black has few options) but black keeps it going for 10/12 moves?

% \bibliography{paper}{}
% \bibliographystyle{plain}
\end{document}

% rnbqkbnr/pppppppp/8/8/8/8/NNNBBNNN/QQQQKQQQ w KQkq - 0 1