### 2015-03-15 portmantout/paper.tex
Makes particles successfully

\documentclass{article}
\usepackage[top=0.4in, left=0.4in, right=0.4in]{geometry}

\usepackage{setspace}
\usepackage{seqsplit}
\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{code}
\usepackage{graphicx}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{textcomp}

\pagestyle{empty}

\newcommand\comment[1]{}

\newcommand\st{$^{\mathrm{st}}$}
\newcommand\nd{$^{\mathrm{nd}}$}
\newcommand\rd{$^{\mathrm{rd}}$}
\renewcommand\th{$^{\mathrm{th}}$}
\newcommand\tm{$^{\mbox{\tiny \textsc{tm}}}$}

\newcommand\sfrac[2]{{}\,$^{#1}$\!/{}\!$_{#2}$}

\newcommand\citef[1]{\addtocounter{footnote}{1}\footnotetext{\cite{#1}}\ensuremath{^{\mbox{\footnotesize [\thefootnote]}}}}
\begin{document}

\title{Portmantout}
\author{Dr.~Tom~Murphy~VII~Ph.D.\thanks{
WIKIPEDIA!!!!
}
}

\renewcommand\>{$>$}
\newcommand\<{$<$}

\date{15 Mar 2015}

\maketitle

\begin{abstract}
Long word ok cool
\end{abstract}

\vspace{1em}
{\noindent \small {\bf Keywords}:
 super
}

\newcommand\foo[1]{%
  \begin{minipage}{7.6in}
  \seqsplit{#1}
  \end{minipage}
  }

\section{Introduction}

A {\em portmanteau}, henceforewith non-italicized, is a
stringin'-together of two words to make a new word, like
``brogrammer'' ({\sf brother} + {\sf programmer}; a programmer who is
your brother), or (other example).

It is natural to think of generalizations of the portmanteau, such as
the {\em portmantrois},\!\footnote{Graham Smith, personal
  communication.} (itself a portmanteau of {\sf portmanteau} + {\sf
  trois}, French-language for three) the human-centipedification of
{\em three} words, such as ``anachillaxis'' ({\sf anaphylaxis} + {\sf
  chill} + {\sf relax}; a severe allergic reaction to idleness) or
``brogrammermaid'' ({\sf brother} + {\sf programmer} + {\sf mermaid};
a programmer who is your brother and a mermaid).

In this paper I present the world's first (?)\footnote{I did a couple
  Google searches; seems good enough.} {\em portmantout}, a
portmanteau of all English-language words.

overmagnifyingearlesshrimpierabbinicalcicadaeratorshrimpiestandardizableaseholdershrimpingementshrimpsychedelically ({\sf overmagnifying} + {\sf gearless} + {\sf shrimpier} + {\sf rabbinical} + {\sf calc} + {\sf cicada} + {\sf aerator} + {\sf shrimpiest} + {\sf standardizable} + {\sf leaseholders} + {\sf shrimping} + {\sf impingements} + {\sf shrimps} + {\sf psychedelically}; presumably having something to do with shrimp).

\section{The big ol' word}
\tiny
\setstretch{0.1}
biiiiig word.

.... put big word here ....

\end{document}