### 2016-03-06 trunk/snacks/paper.tex
More snack sorting policies, notes for paper

\documentclass[twocolumn]{article}
\usepackage[top=1.1in, left=0.85in, right=0.85in]{geometry}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{code}
\usepackage{graphicx}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{textcomp}

\pagestyle{empty}

\newcommand\comment[1]{}

\newcommand\st{$^{\mathrm{st}}$}
\newcommand\nd{$^{\mathrm{nd}}$}
\newcommand\rd{$^{\mathrm{rd}}$}
\renewcommand\th{$^{\mathrm{th}}$}
\newcommand\tm{$^{\mbox{\tiny \textsc{tm}}}$}

% nice fractions
\newcommand\sfrac[2]{{}\,$^{#1}$\!/{}\!$_{#2}$}

\newcommand\citef[1]{\addtocounter{footnote}{1}\footnotetext{\cite{#1}}\ensuremath{^{\mbox{\footnotesize [\thefootnote]}}}}

\usepackage{ulem}
% go back to italics for emphasis, though
\normalem

\begin{document}

\title{Snacking policies}
\author{Dr.~Tom~Murphy~VII~Ph.D.\thanks{
    Copyright \copyright\ 2016 the Regents of the Wikiplia Foundation.
    Appears in SIGBOVIK 2016 with the haunting memory of the
    Association for Computational Heresy; {\em IEEEEEE!} press,
    Verlag--Verlag volume no.~0x2016. ANG 0.00} }

\renewcommand\>{$>$}
\newcommand\<{$<$}

\date{1 April 2016}

\maketitle

\begin{abstract}
% XXX better abstract
what do I put here
\end{abstract}

\vspace{1em}
{\noindent \small {\bf Keywords}:
  snacking, consequentialism, comestability theory
}

\section{Introduction}

Some workplaces of the future

308000 rounds
Without reordering, utils: [589.02, 780.55] mean 684.87

Player 0 always sorts from the shelf he takes:

% 315000 rounds
% utils: [589.57, 792.09] mean 684.96
% p0: [9.11, 36.71] mean 22.59
% ineq: [20.34, 41.60] mean 30.60

total: 684.73 mean, 585.76--781.39
p0: 22.58, 8.82--36.77
ineq: 30.59, 19.93--41.84

No resorting:
total: 685.02 mean, 586.49--781.99
p0: 22.86, 9.50--37.97
ineq: 30.60, 19.92--41.79

Everyone resorts:
total: 684.46 mean, 586.26 -- 780.25
p0: 22.83 mean, 8.96-37.40
ieq: 30.55 19.94--41.27

--------------------

with 0.05 cost to look

total utils: 594.49 mean, 509.82-681.81
p0: 19.83 mean, 7.22-33.24
ieq: 28.07, 18.30-38.49

only p0 sorts:
total utils: 585.52 mean, 499-669
p0: 18.65 7.07-31.26
ieq: 27.92 18.33-38.38

Everyone puts favorite last:
total: 546.61 [466.36, 626.97]
p0: 18.24 [6.39, 31.27]
ieq: 26.92 [17.73, 37.05]

--------------------------------------------------
Now using 100k iterations. everyone puts snack last:

snacks at end × 100000 = Mean: 201.47 [116.62 , 443.72] Median: 208.81
total utils × 100000 = Mean: 546.75 [466.20 , 625.91] Median: 539.81
p0 utils × 100000 = Mean: 18.24 [6.06 , 30.91] Median: 16.52
inequality × 100000 = Mean: 26.91 [17.90 , 37.38] Median: 25.14

nobody sorts:
snacks at end × 100000 = Mean: 201.65 [116.65 , 421.74] Median: 217.99
total utils × 100000 = Mean: 594.17 [508.66 , 680.64] Median: 590.98
p0 utils × 100000 = Mean: 19.82 [7.52 , 33.50] Median: 19.63
inequality × 100000 = Mean: 28.05 [18.23 , 38.42] Median: 26.23

we eat almost exactly the same number of snacks, but are less happy.

nobody sorts; measure snack utils specifically:

snacks at end × 100000 = Mean: 201.26 [116.59 , 417.74] Median: 203.90
total utils × 100000 = Mean: 594.45 [506.68 , 679.40] Median: 588.88
p0 utils × 100000 = Mean: 19.83 [7.53 , 33.60] Median: 17.61
p0 snack utils × 100000 = Mean: 22.30 [8.49 , 36.57] Median: 20.62
inequality × 100000 = Mean: 28.05 [18.31 , 38.37] Median: 26.00

only p0 sorts:

snacks at end × 100000 = Mean: 201.52 [117.71 , 421.74] Median: 198.99
total utils × 100000 = Mean: 590.13 [507.03 , 678.05] Median: 590.58
p0 utils × 100000 = Mean: 18.98 [7.34 , 31.92] Median: 18.44
p0 snack utils × 100000 = Mean: 22.19 [9.17 , 37.02] Median: 20.87
inequality × 100000 = Mean: 27.99 [18.52 , 38.53] Median: 25.96

only p0 places outliers:

snacks at end × 100000 = Mean: 201.54 [117.79, 435.73] Median: 198.50
total utils × 100000 = Mean: 592.27 [509.32, 681.56] Median: 593.03
p0 utils × 100000 = Mean: 19.48 [7.14, 32.58] Median: 17.97
p0 snack utils × 100000 = Mean: 22.28 [8.53, 36.59] Median: 19.70
inequality × 100000 = Mean: 28.01 [18.51, 38.60] Median: 27.01

Credible interval\cite{krushke2014doing}

\end{document}